Making Sequential Programs Deterministic

### 2019-03-22 design.tex
Comment out a section

\section{Design}
\subsection{Type checking background}
The major components of a type system include: 1) \textit{types}, 2) \textit{subtyping rules}, and 3) \textit{dataflow analysis}.
A \textit{type} serves as an abstraction for the set of acceptable values for any expression. The types in a type system form a lattice of finite height. 
The hierarchy of types in this lattice defines subtyping relationships among them.
In our framework, we require every type hierarchy to define a unique\<@Top> and a \<@Bottom> element. This ensures that
any given pair of types has a \textit{least upper bound} and a \textit{greatest lower bound}.
Throughout this paper, we use the notation \<@A> <: \<@B> to denote that type \<@A> is a subtype of type \<@B>.
As an illustration of the type hierarchy, consider the lattice of types shown in Figure~\ref{fig-example-lattice}.
\begin{figure}
    \begin{center}
        \includegraphics[scale=0.15]{lattice}
    \end{center}
    \caption{Example type hierarchy}
    \label{fig-example-lattice}
\end{figure}
In this type system, \<@C> is a subtype of \<@A> as well as \<@B> and types \<@A> and \<@B> are incomparable.
The type checker now performs additional type checking (similar to javac's type checking) with respect to this type hierarchy
and reports any violations. The snippet in Figure~\ref{code-invalid1} would result in an error at the assignment statements:
\begin{figure}
    \begin{verbatim}
    @A int x; @B int y; @C int z;
    x = y;    // Because types @A and @B are incomparable.
    z = x;    // Because @C is a subtype of @A.
    \end{verbatim}
\caption{Example: Invalid assingment}
\label{code-invalid1}
\end{figure}

Similarly, if the return type of a method is annotated as \<@C> but it returns a type that has either \<@A> or \<@B>, the
checker would report an error at compile time.

%Method overriding and example
For method overriding, the usual rules apply for parameters and arguments i.e covariant subtyping for parameters
and contravariant subtyping for return types. Consider the example in Figure~\ref{code-invalid2} where method \<foo> of \<class X> is overridden
in \<class Y>.
\begin{figure}
    \begin{verbatim}
    class X {
        @B int foo(@C int param) {}
    }
    class Y extends X {
        @C int foo(@B int param) {}
    }
    \end{verbatim}
    \caption{Example: Invalid Method override}
    \label{code-invalid2}
\end{figure}

The overriding method is invalid for two reasons: type of the parameter (\<@B>) is not a subtype of the parameter type in the overriden method (\<@C>) and the return type (\<@C>) is not a supertype of the return type of the overriden method (\<@B>).

%collections type parameters - invariant
Collection types are invariant. So, if a \<List> x is declared as \codeid{@A List<@A String> x;} and another \<List y> is declared
as  \codeid{@A List<@B String> y;}, writing \<x = y> would result in an error.

%covariant array types
Arrays follow covariant subtyping rules in java. Therefore, the assignment statement \<@A int @A[] x = y;> where \<y> was 
declared as \<@B int @A[] y> would type check without any warning.

>>>>

\section{Design}

%% This section is too basic, and therefore is boring because most readers
%% will already know it.  It also delays the interesting parts of the paper
%% for too long.  We can introduce concepts as they are needed in the rest
%% of the paper.
% \subsection{Type checking background}
% The major components of a type system include: 1) \textit{types}, 2) \textit{subtyping rules}, and 3) \textit{dataflow analysis}.
% A \textit{type} serves as an abstraction for the set of acceptable values for any expression. The types in a type system form a lattice of finite height. 
% The hierarchy of types in this lattice defines subtyping relationships among them.
% In our framework, we require every type hierarchy to define a unique\<@Top> and a \<@Bottom> element. This ensures that
% any given pair of types has a \textit{least upper bound} and a
% \textit{greatest lower bound}.\todo{It guarantees that any pair of types
%   has an upper and a lower bound, but not that a unique least upper bound
%   and greatest lower bound exist.}
% Throughout this paper, we use the notation \<@A> <: \<@B> to denote that type \<@A> is a subtype of type \<@B>.
% As an illustration of the type hierarchy, consider the lattice of types shown in Figure~\ref{fig-example-lattice}.
% \begin{figure}
%     \begin{center}
%         \includegraphics[scale=0.15]{lattice}
%     \end{center}
%     \caption{Example type hierarchy}
%     \label{fig-example-lattice}
% \end{figure}
% In this type system, \<@C> is a subtype of \<@A> as well as \<@B> and types \<@A> and \<@B> are incomparable.
% The type checker now performs additional type checking (similar to javac's type checking) with respect to this type hierarchy
% and reports any violations. The snippet in Figure~\ref{code-invalid1} would result in an error at the assignment statements:
% \begin{figure}
%     \begin{verbatim}
%     @A int x; @B int y; @C int z;
%     x = y;    // Because types @A and @B are incomparable.
%     z = x;    // Because @C is a subtype of @A.
%     \end{verbatim}
% \caption{Example: Invalid assingment}
% \label{code-invalid1}
% \end{figure}
% 
% Similarly, if the return type of a method is annotated as \<@C> but it returns a type that has either \<@A> or \<@B>, the
% checker would report an error at compile time.
% 
% %Method overriding and example
% For method overriding, the usual rules apply for parameters and arguments i.e covariant subtyping for parameters
% and contravariant subtyping for return types. Consider the example in Figure~\ref{code-invalid2} where method \<foo> of \<class X> is overridden
% in \<class Y>.
% \begin{figure}
%     \begin{verbatim}
%     class X {
%         @B int foo(@C int param) {}
%     }
%     class Y extends X {
%         @C int foo(@B int param) {}
%     }
%     \end{verbatim}
%     \caption{Example: Invalid Method override}
%     \label{code-invalid2}
% \end{figure}
% 
% The overriding method is invalid for two reasons: type of the parameter (\<@B>) is not a subtype of the parameter type in the overriden method (\<@C>) and the return type (\<@C>) is not a supertype of the return type of the overriden method (\<@B>).
% 
% %collections type parameters - invariant
% Collection types are invariant. So, if a \<List> x is declared as \codeid{@A List<@A String> x;} and another \<List y> is declared
% as  \codeid{@A List<@B String> y;}, writing \<x = y> would result in an error.
% 
% %covariant array types
% Arrays follow covariant subtyping rules in java. Therefore, the assignment statement \<@A int @A[] x = y;> where \<y> was 
% declared as \<@B int @A[] y> would type check without any warning.
